---
title: "Exercise 1"
format: html
editor: visual
---

```{r}
# First run this
install.packages("pacman")

```

### Libraries

```{r}
library(pacman)

p_load(dlookr,
       DMwR2, # Data Mining with R functions
       GGally, # Pair-wise plots using ggplot2
       Hmisc, # Data analysis 
       palmerpenguins, # Alternative to the Iris dataset
       tidyverse) # Data wrangling, manipulation, visualization
```

### Loading the Data

```{r}
data(algae, package = "DMwR2")

algae |> glimpse()    
```

### Central Tendencies: Mean, Median & Mode

### Mean

```{r}
algae$a1 |>
  mean()      #Returns mean of col a1

```

### Median

```{r}
algae$a1 |>
  median()    #Returns median of col a1

```

### Mode

```{r}
#Function for Mode

Mode <- function(x, na.rm=FALSE){
if(na.rm) x<-x[!is.na(x)] #Checking the similarity
ux <- unique (x) #Storing unique values in ux
return (ux[which.max(tabulate(match(x, ux)))]) #Returns the freq of each of the unique values
}

algae$a2 |> Mode()

```

### DMwR centralValue() function:

```{r}
# Numerical variable
algae$a1 |> centralValue() #returns median for numeric
```

```{r}
# Numerical variable
algae$speed |> centralValue() #returns mode for nominal

```

### Statistics of spread

### Variance

```{r}
algae$a1 |> var() # returns variance of col a1
```

### Standard deviation

```{r}
algae$a1 |> sd() # returns standard deviation of col a1
```

### Range

```{r}
algae$a1 |> range() # returns teh range of the col a1
```

### Maximum Value

```{r}
algae$a1 |> max() # returns the maximum value in a1
```

### Minimum Value

```{r}
algae$a1 |> min() # returns the minimum value in a
```

### Interquartile

```{r}
algae$a1 |> IQR() # returns the IQR of a
```

### Quantiles

```{r}
algae$a1 |> quantile() # returns quantils of a1
```

```{r}
algae$a1 |> quantile(probs = c(0.2, 0.8)) # specific quantiles
```

### Missing values

```{r}
library(purrr)
# Computes the total number of NA values in the dataset
nas <- algae %>%                              
  purrr::map_dbl(~sum(is.na(.))) %>% 
  sum()

cat("The dataset contains ", nas, "NA values. \n")
```

```{r}
# Computes the number of incomplete rows in the dataset
incomplete_rows <- algae %>% 
  summarise_all(~!complete.cases(.)) %>%
  nrow()
```

```{r}
cat("The dataset contains ", incomplete_rows, "(out of ", nrow(algae),") incomplete rows. \n")

```

### Summaries of a Dataset

```{r}
algae |> summary()
```

### Hmisc's describe()

```{r}
data("penguins")
penguins |> Hmisc::describe()
```

### dlookr's describe()

```{r}
penguins |> dlookr::describe()
```

### Summaries on a subset of data

```{r}
algae |>
  summarise(avgNO3 = mean(NO3, na.rm=TRUE),
            medA1 = median(a1))
```

```{r}
algae |>
  select(mxPH:Cl) |>
  summarise_all(list(mean, median), na.rm = TRUE)
```

```{r}
algae |>
  select(a1:a7) |>
  summarise_all(funs(var))

```

```{r}
algae |>
  select(a1:a7) |>
  summarise_all(c("min", "max"))
```

### Use summarise() with group_by()

```{r}
algae |>
  group_by(season, size) |>
  summarise(nObs = n(), mA7 = median(a7))
```

```{r}
penguins |> 
  group_by(species) |>
  summarise(var = var(bill_length_mm, na.rm = TRUE))
```

### Aggregating data

```{r}
#Can be helpful for summary functions that don’t return a scalar

penguins |>
  group_by(species) |>
  reframe(var = quantile(bill_length_mm, na.rm = TRUE))
```

```{r}
penguins |>
  group_by(species) |>
  dlookr::describe(bill_length_mm)
```
